### 2015-02-27 rmtw/a5/latex8.tex
Review protocol d0ne

%
%  $Description: Author guidelines and sample document in LaTeX 2.09$ 
%
%  $Author: John Burchell & William Granli $
%  $Date: 2015/01/21 15:20:59 $
%  $Revision: 1.0
%

\documentclass[10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{url}
\usepackage{tabularx}
\usepackage{appendix}

%\documentstyle[times,art10,twocolumn,latex8]{article}

%------------------------------------------------------------------------- 
% take the % away on next line to produce the final camera-ready version 
\pagestyle{empty}

%------------------------------------------------------------------------- 
\begin{document}

\title{Assignment 5}

% author names and affiliations
\author{John Burchell and William Granli \\
john.a.burchell@gmail.com, william.granli@gmail.com}

\maketitle
\thispagestyle{empty}

%-------------------------------------------------------------------------

​\Section{Introduction}
Ever since the first computer programs were written, researchers and developers have been working on new programming concepts in order to decrease the level of complexity needed to develop programs. The abstraction level has been increased with help of compilers, object-oriented programming and Unified Modelling Language (UML). The current trend within software engineering (SE) is Model-driven Software Development (MDSD) which uses UML to generate object-oriented code \cite{staron2006adopting}.

The benefits of MDSD have consistently been reported in academia \cite{staron2006adopting} \cite{volter2013model}. Many large companies such as Ericsson and ABB \cite{staron2006adopting} also use MDSD as a part of their development chain. In industry, however, MDSD has struggled to make a complete breakthrough and it is mostly used for niche systems such as embedded and safety-critical systems.

\SubSection{Statement of the Problem}

The selection of development process, programming language and programming paradigm has a high impact on the development speed depending on what the the system requirements are and in which environment the system is developed. Selecting for example MDSD for the wrong system and setting may have a very negative impact on development speed, quality and maintainability.

\SubSection{Existing Literature}

Volter et al\cite{volter2013model} says that successful adoption of MDSD should increase development speed, increase software quality, improve reusability and improve manageability through abstraction.

Whittle and Hutchinson \cite{whittle2012mismatches} conducted a study whereby they examined 17 companies who were adopting MDSD. While they agreed that the successful adoption of MDSD can improve productivity through code generation, there are many pitfalls which can cause the adoption to fail. Whittle and Hutchinson found that in companies where the change was implemented bottom up, there was a higher degree of success. On the contrary to this, when change was top down, as suggested by academia and education, the adoption was often not as successful. A focus on Domain Specific Languages (DSLs) is what Whittle and Hutchinson suggest helps MDSD become successful in a company setting.

A further, more specific example is a study conducted by Staron \cite{staron2006adopting}. In this study a case study is conducted with two companies wishing or in the process of adopting MDSD. In the findings, Staron provides a set of conditions for adopting MDSD, one being the compatibility of MDSD and existing principles in the company. Staron states that the adoption of this new process must be usable by the previous processes in place \cite{staron2006adopting}. He concludes by stating that this compatibility is especially important for larger companies due to the heavy effort required to implement the change. In a second study \cite{staron2008transitioning}, Staron provides possible reasons for the adoption of MDSD to fail. He suggests that the large investment in time and money, immature technology and organisational resistance are all factors to overcome in the adoption of new processes.

Mattsson et al. \cite{mattsson2009linking} discuss the correlation between MDSD and software architecture. The idea that MDSD is beneficial if performed correctly is presented, and one important factor discussed is the need of automatic enforcement of architectural rules \cite{mattsson2009linking}. If this has to be performed manually, it will become a bottleneck in large MDSD projects \cite{mattsson2009linking}.

\SubSection{Motivation for Study}

Academia appears to be united in the opinion that MDSD is an improvement compared with traditional software development \cite{staron2006adopting} \cite{volter2013model}. However, why do many large software companies chose not to adopt MDSD?

This gap in the literature is the motivation for our study. There is therefore a need to methodically investigate the existing literature in which studies of the adoption of MDSD, advantages and disadvantages of MDSD are explored. Not only this but how and when should MDSD be adopted and how it can be ensured to succeed.

\SubSection{Research Questions}
\begin{itemize}
\item \textbf{RQ1:} In which situations should companies expect that changing to MDSD would lead to positive effects? 
\item \textbf{RQ1.1:} In which domain(s) is MDSD the most successful? 
\item \textbf{RQ1.2:} Which software development process(es) has been most successfully used in conjunction with MDSD?
\item \textbf{RQ1.3:} What are the characteristics of the systems of the companies that successfully use MDSD?
\item \textbf{RQ1.4:} Which change strategy has most successful in implementing MDSD? 
\end{itemize}

\Section{Methodology}
The method used to attempt to answer our research questions was a Systematic Literature Review (SLR) \cite{keele2007guidelines}. An SLR was the most appropriate research methodology for two main reasons. Firstly, we wished to summarise the existing evidence of when MDSD is applicable in industry and secondly to fill gaps in literature for examples of when MDSD is not applicable \cite{keele2007guidelines}. For these same reasons, other methodology techniques were not as applicable. For example, a quantitative study was not appropriate as we were not trying to prove, disprove or generalise a hypothesis. A qualitative study would have been useful if we were interested in exploring a hypothesis. For similar reasons a case study would also not have been applicable as we were not concerning ourselves with how MDSD works in a real setting. Likewise, as industry and academia seem to differ in opinion, there is an implication that there could have been be bias in both sides of the argument. By performing an SLR we helped to alleviate some of this bias but we did not eliminate it entirely \cite{keele2007guidelines}.

It is important to note that did not perform either a Systematic Mapping Study nor a Tertiary Review of existing literature. There was sufficient evidence in existence for the applicability of MDSD such that we could rule out the need to perform a Systematic Mapping Study. Furthermore, there were insufficient SLRs that had been conducted prior to our study, thus, a Tertiary Review was not possible nor appropriate at that time.

\SubSection{Planning the Study}
Initially, existing SLRs were identified and evaluated, with the goal of mapping and identifying the gap in the existing literature. All identified SLRs \cite{santiago2012model} \cite{dias2007survey} \cite{loniewski2010systematic} \cite{nicolas2009generation} were, however, researching the usage of model-driven techniques in other fields, such as traceability management \cite{santiago2012model}, testing \cite{dias2007survey} and requirements engineering \cite{loniewski2010systematic} \cite{nicolas2009generation}. Based on these results, the need to evaluate previous research on model driven \textit{software} development was identified.

The next step was to define the research questions. The main research question was constructed to create a wide scope to be able to investigate which characteristics are most vital to consider when adopting MDSD. The sub-questions were constructed to provide more detailed answers for specific questions that are valuable to the industry and companies who are considering to implement MDSD.

\subsubsection{Review Protocol}
The next step was to (<- XXX re-phrase that) define the review protocol. The list of databases included can be seen in Table I. The included databases have been chosen based on recommendations in \cite{brereton2007lessons}. SpringerLink was included as advised in \cite{keele2007guidelines}. The databases were divided evenly between the two researchers.

The search terms used can be found in Appendix A. The main search terms are related to MDSD, but related terms such as Model Driven Engineering and code generation were also included.

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|} 
		\hline
		\textbf{Database} & \textbf{Link}  \\
		\hline
		IEEExplore &  Granli \\
		\hline
		ACM Digital library &  Granli \\
		\hline
		Google Scholar &  Granli \\
		\hline
		Citeseer Library & Granli \\
		\hline
		Inspec &  Burchell \\
		\hline
		ScienceDirect & Burchell \\
		\hline
		EI Compendex &  Burchell \\
		\hline
		SpringerLink & Burchell \\
		\hline
	\end{tabular}
	\caption{Databases included in review}
\end{table}

The inclusion criteria were defined as follows: \newline
\begin{itemize} 
\item Studies including examples about the application of MDSD in industry
\end{itemize}

The exclusion critiera were defined as follows: \newline
\begin{itemize}
\item Informal papers (no defined research questions, no methodology)
\item Non-peer reviewed papers (books, white papers etc.)
\end{itemize}

Once papers were selected, the following data was selected: Publication year, topic area, methodology type, authors, summary of the paper, conclusion. This data was the used as a base for analysis when evaluating the research questions for each paper.

Evaluating the review protocol (5.5)

\SubSection{Conducting}
\textbf{Conducting the review}
\begin{itemize}
\item Identification of research (6.1)

Trial Search, Checking vs already performed studies.
We picked 6 well known studies, our search strings did manage to find them all.

Didn't check other disciplines as the area is not relevant.

Summarise publication bias.

Table, showing what sources we searched, what dates, what strategy, Years covered by search, what journal and conference proceedings were covered.

Justify the when, the where and the how.

\item Selection of primary studies (6.2)
Sources that provide direct evidence about the research question. Refined search criteria here?

\item Study quality assessment (6.3) x

Apply the inclusion and exclusion criteria here, keeping a list of all excluded articles.

Studies not related to Software Engineering were excluded first and as such have not been included in the exclusion list due to their irrelevance.

Cohen Kappa statistic. Initial value here. Disagreements will be discussed and resolved.

\item Data extraction and monitoring (6.4) x

Quality of the papers must also be considered before they are examined for use in the study, this is done along with the inclusion and exclusion criteria.

\item Data synthesis x
\end{itemize}

\SubSection{Reporting (probably skip)}
\textbf{Reporting the review}
\begin{itemize}
\item Specifying dissemination mechanisms (7.1)
\item Formatting the main report (7.2)
\item Evaluating the report (7.3)
\end{itemize}

\bibliographystyle{latex8}
\bibliography{latex8}

\Section{Acknowledgements}

\begin{itemize}
\item \textbf{these bulletins are from the assignment specification}
\item Describe and explain the research strategy (i.e. SLR) and investigation methods that will be used for the study. 
\item Motivate your choice of experiment using relevant references. 
\item Search string including keywords and used databases
\item inclusion/exclusion criteria
\item data extraction
\item data synthesis
\item quality assessment
\end{itemize}

\begin{appendices}
\section{Search Terms}
MDD, MDSD, model-driven software development, model-driven development, MDE, model-driven engineering, model driven engineering, code generation
\end{appendices}

\end{document}